Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-24 paper.tex
revise sec2

The dynamic semantics of a cast calculus is typically defined with a
reduction semantics. Here we use a CEK machine
\citep{felleisen1986control} instead because the first author is more
familiar with CEK machines and believes that a CEK machine is more
convenient to use for the space-efficient semantics in
Section~\ref{sec:framework:cek} (a major point of space-efficiency is about 
compressing continuations). So using an abstract machine for the cast calculi in
this section makes it easier prove correctness of the space-efficient
machines. Of course, one should prove that the abstract machine
presented here is equivalent to the standard reduction semantics for
cast calculi, but we have not yet done so.

Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
machine and Fig.~\ref{fig:state} gives a grammar for machine states
$s$, including a definition of values and value typing. 
%Fig.~\ref{fig:state} gives a grammar for machine states
%$s$, including a definition of values and value typing and
%Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
%machine.
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machine involves two kinds of transitions, (1) those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation, and (2) those that return a value to the
current continuation and possibly perform a computation.
Corresponding to (1) and (2), the machine state is either in an
$\sOOinspect{e}{\rho}{\kappa}$ or $\sOOreturn{v}{\kappa}$ configuration, 
respectively. Additionally, there is the $\sOOhalt{o}$ configuration, where the 
machine halts with an observation.
%Let $s$ range over machine states. A state is either looking at an 
%expression to decide what to do next, returning a value to a continuation, or 
%halting with an observation.

@@

The transition relation $\judgeCreduce{s}{s}$ is parameterized over
$applyCast$ to allow for the differences between D and UD.
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%% Other
%% transition rules starting from$Eval$states are standard.
To apply a function in cast, the machine first casts $v_1$, the
operand, then applies the cast operand to $v_2$, the underlying
function, and finally cast the return value.
%
To take out the first (resp. second) part of a pair in cast, the
machine firstly take out the first (resp. second) part of $v$, the
underlying pair, and cast the result.
%
%In case splitting, if the target value is a left injection, the
%machine moves to a state that will apply the first continuation
%function to the value inside the left injection. The case for right
%injection is similar.
%
To case split an injection in cast, the machine moves the cast from
the injection to continuations functions.
%
To cast a value, the machine invokes $applyCast$ on the value. If the cast 
succeeds ($\rOOsucc{v'}$), the machine returns the result to the next
continuation.  If the cast fails ($\rOOfail{l}$), the machine halts with the 
blame label.
%Let $r$ range over cast results. A cast result is either a success, which 
%brings back a value, or a failure, which brings back a blame label.

@@

The reflexive transitive closure of reduction ($\judgeCreduceTrans{s}{s}$) and 
evaluation ($eval_\mathcal{x}(e)$) are standard~\citep{felleisen03:_pllc}.
Let $o$ range over observations. They are what we would observe if 
a program terminates. Observations include all value constructors and
blame.  The function converting values to observations ($observe(v) =
o$) is defined in the natural way.

>>>>

The dynamic semantics of a cast calculus is typically defined with a
reduction semantics. Here we use a CEK machine
\citep{felleisen1986control} instead because the first author is more
familiar with CEK machines and believes that a CEK machine is more
convenient to use for the space-efficient semantics in
Section~\ref{sec:framework:cek} (a major point of space-efficiency is about 
compressing continuations). So using an abstract machine for the cast calculi in
this section makes it easier to prove correctness of the space-efficient
machines. Of course, one should prove that the abstract machine
presented here is equivalent to the standard reduction semantics for
cast calculi, but we have not yet done so.

Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
machine and Fig.~\ref{fig:state} gives a grammar for machine states
$s$, including a definition of values $v$ and value typing. 
%Fig.~\ref{fig:state} gives a grammar for machine states
%$s$, including a definition of values and value typing and
%Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
%machine.
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machine involves two kinds of transitions, (1) those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation, and (2) those that return a value to the
current continuation and possibly perform a computation.
Corresponding to (1) and (2), the machine state is either in an
$\sOOinspect{e}{\rho}{\kappa}$ or $\sOOreturn{v}{\kappa}$ configuration, 
respectively. Additionally, there is the $\sOOhalt{o}$ configuration, where the 
machine halts with an observation ($o$). Observations include all value 
constructors and blame. The function converting values to 
observations ($observe(v) = o$) is defined in the natural way.
%Let $s$ range over machine states. A state is either looking at an 
%expression to decide what to do next, returning a value to a continuation, or 
%halting with an observation.

@@

The transition relation $\judgeCreduce{s}{s}$ is parameterized over
$applyCast$ to allow for the differences between D and UD.
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%% Other
%% transition rules starting from$Eval$states are standard.
To apply a casted function, the machine first casts $v_1$, the
operand, then applies the cast operand to $v_2$, the underlying
function, and finally cast the return value.
%
To take out the first (resp. second) part of a casted pair, the
machine firstly take out the first (resp. second) part of $v$, the
underlying pair, and cast the result.
%
%In case splitting, if the target value is a left injection, the
%machine moves to a state that will apply the first continuation
%function to the value inside the left injection. The case for right
%injection is similar.
%
To case split a casted injection, the machine moves the cast from
the injection to continuations functions.
%
To cast a value, the machine invokes $applyCast$ on the value. If the cast 
succeeds ($\rOOsucc{v'}$), the machine returns the result to the next
continuation.  If the cast fails ($\rOOfail{l}$), the machine halts with the 
blame label.
%Let $r$ range over cast results. A cast result is either a success, which 
%brings back a value, or a failure, which brings back a blame label.

@@

The reflexive transitive closure of reduction ($\judgeCreduceTrans{s}{s}$) and 
evaluation ($eval_\mathcal{x}(e)$) are standard~\citep{felleisen03:_pllc}.
